\documentclass{article}

\usepackage{
	arxiv,
	amsmath,
	amssymb,
	cleveref
}

\title{Computing the Metric Dimension of Graphs using Satisfiability Modulo Theories}

\author{
	Fikri Yudhistira Ajiwijaya \\
	Institut Teknologi Sepuluh Nopember
	\And
	Muhammad Syifa'ul Mufid \\
	Institut Teknologi Sepuluh Nopember
}

\begin{document}
	\maketitle

	\begin{abstract}
		...\textit{blah blah blah}...
	\end{abstract}

	\section{Introduction}

	The theoretical part of this paper is divided into three major parts. The first of which consists of the heavy lifting of formally defining notations for slicing, dicing, and filtering of matrices, tuples, and sets. While it may seem trivial and verbose at first glance, it is of utmost importance due to the lack of standard notations for these basic operations. Part two and part three extensively relies on using these carefully defined operations, without which would be incomprehensible to read. This rigor is both necessary and unnegotiable. Part two consists of a quick recall on the metric dimension problem definition, then followed by the first major contribution of this paper; that is, the journey in perspective shift of metric dimension reformulated in terms of slicing, dicing, and filtering. After all this gruelling groundwork, we are finally able to introduce the theoretical part of our result in part three, where we formulate constraints for metric dimension in satisfiability modulo theories, through careful solution space analysis.

	\section{Slicing, Dicing, and Filtering}

	For any set \(U\), we denote its cardinality as \(|U|\) and its power set as \(\mathcal{P}\). Given a subset \(R \subseteq U\), its complement is defined as \(\widetilde{R} = U \backslash R\).

	For a \(k\)-tuple \(C\), we denote its length by \(|C|\) and refer to the element at the \(i\)-th position as \(C(i)\). We say that \(U\) has a correspondence with \(C\) when there exists a bijective function \(\mathrm{idx} : U \to \{1, \cdots, |C|\}\). This allows us to index \(C\) by elements of \(U\), where for any \(u \in U\), we define \(C(u) = C(\mathrm{idx}(u))\). Furthermore, for any subset \(\{u_1, \cdots, u_r\} \subseteq U\), we define the sub-tuple \(C(\{u_1, \cdots, u_r\}) = (C(u_1),\ \cdots,\ C(u_r))\), preserving element order of \(C\).

	We define some element-wise operations on \(k\)-tuples. The element-wise negation, \(\neg C = (\neg C(1), \cdots, \neg C(k))\). The element-wise disjunction, \(C_1 \vee C_2 = (C_1(1) \vee C_2(1), \cdots, C_1(k) \vee C_2(k))\). The element-wise equality, \(C_1\ \ddot{=}\ C_2 = (C_1(1) = C_2(1), \cdots, C_1(k) = C_2(k))\). We also define \((\mathrm{True})^k\) as the \(k\)-tuple consisting of all \(\mathrm{True}\) values.

	Given a Boolean \(k\)-tuple \(B\), where \(U\) has a correspondence with \(B\), we define the filter operation \(U(B) = \{u \in U : B(u) = \mathrm{True}\}\). Similarly, given an integer \(k\)-tuple \(I\) that \(U\) has a correspondence with, \(U(I) = \{u \in U : I(u) \neq 0\}\). We extend this operation for a set of Boolean \(k\)-tuples, \(U(\{B_1, B_2, \cdots\}) = \{U(B_1),\ U(B_2),\ \cdots\}\). Likewise, for a set of integer \(k\)-tuples, \(U(\{I_1, I_2, \cdots\}) = \{U(I_1),\ U(I_2),\ \cdots\}\).

	For a matrix \(A\) with \(n\) rows and \(m\) columns, we use \(|A|_{\mathrm{row}}\) to denote the number of rows in \(A\) and \(|A|_{\mathrm{col}}\) to denote the number of columns in \(A\). We define the notation \(\mathcal{A} \overset{\mathrm{row}}{\in} A\) to say that \(\mathcal{A}\) is a row in \(A\) taken as an \(m\)-tuple, and \(\mathcal{A} \overset{\mathrm{col}}{\in} A\) to say that \(\mathcal{A}\) is a column in \(A\) as taken an \(n\)-tuple. We refer to the element at the \(i\)-th row and \(j\)-th column with \(A(i, j)\), the \(i\)-th row taken as an \(m\)-tuple with \(A(i, \cdot)\), and the \(j\)-th column taken as an \(n\)-tuple with \(A(\cdot, j)\).

	We say that a set \(S\) has a row correspondence with \(A\) when there exists a bijective function \(\mathrm{idx} : S \to \{1, \cdots, |A|_{\mathrm{row}}\}\), and say that a set \(T\) has a column correspondence with \(A\) when there exists a bijective function \(\mathrm{idx} : T \to \{1, \cdots, |A|_{\mathrm{col}}\}\). This allows us to index the rows of \(A\) by elements of \(S\) and the columns of \(A\) by elements of \(T\), where for any \(s \in S\), we define \(A(s, \cdot) = A(\mathrm{idx}(s), \cdot)\), and for any \(t \in T\), we define \(A(\cdot, t) = A(\cdot, \mathrm{idx}(t))\).

	Finally, for any subset \(\{s_1, \cdots, s_r\} \subseteq S\), we define the submatrix by row selection as
	\[
		A(\{s_1, \cdots, s_r\}, \cdot) = \left[
		\begin{array}{c}
			A(s_1, \cdot) \\
			\vdots \\
			A(s_r, \cdot)
		\end{array}
		\right],
	\]
	preserving row order of \(A\). And for any subset \(\{t_1, \cdots, t_r\} \subseteq T\), we define the submatrix by column selection, as
	\(
		A(\cdot, \{t_1, \cdots, t_r\} \subseteq T) = \left[\enspace A(\cdot, t_1)\enspace \cdots\enspace A(\cdot, t_r)\enspace \right]
	\),
	preserving column order of \(A\).

	\section{Metric Dimension}

	Let \(G\) be a connected graph with \(n\) vertices \(V = \{v_1, \cdots, v_n\}\) and \(m\) edges \(E = \{e_1, \cdots, e_m\}\). There exists a function \(d(u, v)\), for any \(u, v \in V\), that gives the distance from \(u\) to \(v\). Then, for any \(u, v, w \in V\) and \(e \in E\), if \(e\) is the edge that connects \(v\) and \(w\), it is defined \(d(u, e) = \mathrm{min}(d(u, v),\ d(u, w))\). For the reset of this paper, let us take a subset of \(k\) vertices \(W = \{w_1, \cdots, w_k\} \subseteq V\).

	The representation of \(v \in V\) with respect to \(W\) is the \(k\)-tuple \(r(v|W)\), and the edge representation of \(e \in E\) with respect to \(W\) is the \(k\)-tuple \(r(e|W)\), defined as follow
	\begin{equation}
		\label{eq:repr-v-canon}
		r(v|W) = (d(v, w_1), \cdots, d(v, w_k)) ,
	\end{equation}
	\begin{equation}
		\label{eq:repr-e-canon}
		r(e|W) = (d(e, w_1), \cdots, d(e, w_k)) .
	\end{equation}

	\(W\) is said to be a resolving set when all representations with respect to it are unique. And \(W\) is said to be an edge resolving set when all edge representations with respect to it are unique.
	\begin{equation}
		\label{eq:resl-v-canon}
		W \text{ is a resolving set} \iff \forall u, v \in V, u \neq v,\enspace r(u|W) \neq r(v|W)
	\end{equation}
	\begin{equation}
		\label{eq:resl-e-canon}
		W \text{ is an edge resolving set} \iff \forall e, f \in E, e \neq f,\enspace r(e|W) \neq r(f|W)
	\end{equation}

	Conversely, \(W\) is said to be a non-resolving set when there exists a pair of non-unique representations. And \(W\) is said to be an edge non-resolving set when there exists a pair of non-unique edge representations.
	\begin{equation*}
		W \text{ is a non-resolving set} \iff \exists u, v \in V, u \neq v,\enspace r(u|W) = r(v|W)
	\end{equation*}
	\begin{equation*}
		W \text{ is an edge non-resolving set} \iff \exists e, f \in E, e \neq f,\enspace r(e|W) = r(f|W)
	\end{equation*}

	The power set \(\mathcal{P}(V)\) is therefore partitioned into the set of resolving sets and the set of non-resolving sets. Likewise, it is also partitioned into the set of edge resolving sets and the set of edge non-resolving sets.

	Both resolving sets and edge resolving sets has the property of upward monotonicity.
	\begin{equation*}
		W \text{ is a resolving set},\enspace W \subseteq Z \implies Z \text{ is a resolving set}
	\end{equation*}
	\begin{equation*}
		W \text{ is an edge resolving set},\enspace W \subseteq Z \implies Z \text{ is an edge resolving set}
	\end{equation*}

	Both non-resolving sets and edge resolving sets has the property of downward monotonicity.
	\begin{equation*}
		W \text{ is a non-resolving set},\enspace Z \subseteq W \implies Z \text{ is a non-resolving set}
	\end{equation*}
	\begin{equation*}
		W \text{ is an edge non-resolving set},\enspace Z \subseteq W \implies Z \text{ is an edge non-resolving set}
	\end{equation*}

	The metric dimension of \(G\) is the minimum cardinality of its resolving sets, and the edge metric dimension of \(G\) is the minimum cardinality of its edge resolving sets.

	\section{Distance Matrix}

	Let us define the representation matrix and the edge representation matrix with respect to \(W\).
	\begin{equation}
		\label{eq:repr-VW}
		r(V|W) = \left[\enspace r(v_1|W)\enspace \cdots\enspace r(v_n|W)\enspace \right]
	\end{equation}
	\begin{equation}
		\label{eq:repr-EW}
		r(E|W) = \left[\enspace r(e_1|W)\enspace \cdots\enspace r(e_m|W)\enspace \right]
	\end{equation}
	The matrix \(r(V|W)\) has a dimension of \(|W| \times |V|\), and \(V\) has a column correspondence with it. The matrix \(r(E|W)\) has a dimension of \(|W| \times |E|\), and \(E\) has a column correspondence with it.

	By construction, the columns of the representation matrix contain all representations with respect to \(W\). Similarly, the columns of the edge representation matrix contain all edge representations with respect to \(W\). Consequently, this allows us to rewrite \cref{eq:repr-v-canon} and \cref{eq:repr-e-canon} to select a column from these matrices.
	\begin{equation}
		\label{eq:repr-VWv}
		r(v|W) = r(V|W)(\cdot, v)
	\end{equation}
	\begin{equation}
		\label{eq:repr-EWe}
		r(e|W) = r(E|W)(\cdot, e)
	\end{equation}

	We are now able to rewrite \cref{eq:resl-v-canon} to use \cref{eq:repr-VWv}, and \cref{eq:resl-e-canon} to use \cref{eq:repr-EWe}.
	\begin{equation}
		\label{eq:resl-VWv}
		W \text{ is a resolving set} \iff \forall u, v \in V, u \neq v,\enspace r(V|W)(\cdot, u) \neq r(V|W)(\cdot, v)
	\end{equation}
	\begin{equation}
		\label{eq:resl-EWe}
		W \text{ is an edge resolving set} \iff \forall e, f \in E, e \neq f,\enspace r(E|W)(\cdot, e) \neq r(E|W)(\cdot, f)
	\end{equation}
	This has transformed determining that \(W\) is a resolving set as a matter of showing all columns of the representation matrix with respect to \(W\) are unique. Similarly, determining that \(W\) is an edge resolving set has become a matter of showing all columns of the edge representation matrix with respect to \(W\) are unique.

	Let's now shift our focus a little bit. Observe that \(V\) has correspondence with both \(r(v|V)\) and \(r(e|V)\). This mean we can rewrite \cref{eq:repr-v-canon} and \cref{eq:repr-e-canon} to select \(W\) from representation and edge representation with respect to \(V\).
	\begin{equation}
		\label{eq:repr-vVW}
		r(v|W) = r(v|V)(W)
	\end{equation}
	\begin{equation}
		\label{eq:repr-eVW}
		r(e|W) = r(e|V)(W)
	\end{equation}

	Using \cref{eq:repr-vVW} for factorization, we may rewrite \cref{eq:repr-VW} to be a submatrix of \(r(V|V)\) by row selection. Likewise, with \cref{eq:repr-eVW}, we may factorize \cref{eq:repr-EW} to be a submatrix of \(r(E|V)\) by row selection.
	\begin{equation*}
		r(V|W) = r(V|V)(W, \cdot)
	\end{equation*}
	\begin{equation*}
		r(E|W) = r(E|V)(W, \cdot)
	\end{equation*}

	Both \(r(V|V)\) and \(r(E|V)\) are more commonly known as the distance matrix and the edge distance matrix of \(G\), where by we can obtain any distance in \(G\) from either of these matrices.
	\begin{equation*}
		\forall u, v \in V,\enspace d(u, v) = r(V|V)(\cdot, u)(v)
	\end{equation*}
	\begin{equation*}
		\forall v \in V, \forall e \in E, \enspace d(v, e) = r(E|V)(\cdot, v)(e)
	\end{equation*}

	\section{Distance Similarity Matrix}

	Let us now shift our focus to the representation comparison process. For any \(u, v \in V\) and any \(W \subseteq V\), the representation \(r(u|W)\) is equal to the representation \(r(v|W)\) if and only if the tuple \(r(u|W)\ \ddot{=}\ r(v|W)\) doesn't contains all \(\mathrm{True}\); or more formaly, \(r(u|W) = r(v|W) \iff (r(u|W)\ \ddot{=}\ r(v|W)) \neq (\mathrm{True})^{|W|}\). If we take the approach of selecting by \(u, v \in V\) first, this becomes \(r(u|V)(W) = r(v|V)(W) \iff (r(u|V)(W)\ \ddot{=}\ r(v|V)(W)) \neq (\mathrm{True})^{|W|}\). Take a note of the tuple element-wise comparison at the right hand side; we may apply the following factorization: \((r(u|V)(W)\ \ddot{=}\ r(v|V)(W)) = (r(u|V)\ \ddot{=}\ r(v|V))(W)\). This simplifies the representation comparison to \(r(u|V)(W) = r(v|V)(W) \iff (r(u|V)\ \ddot{=}\ r(v|V))(W) \neq (\mathrm{True})^{|W|}\). Finally, we can transitively obtain a new form of the resolving set criteria, to be \(\forall u, v \in V, u \neq v, (r(u|V)\ \ddot{=}\ r(v|V))(W) \neq (\mathrm{True})^{|W|}\).

	When \(W = V\), the resolving set criteria further simplies to \(\forall u, v \in V, u \neq v, (r(u|V)\ \ddot{=}\ r(v|V)) \neq (\mathrm{True})^{|V|}\). By using the distance matrix notation, this becomes \(\forall u, v \in V, u \neq v, (D^{(V)}(\cdot, u)\ \ddot{=}\ D^{(V)}(\cdot, v)) \neq (\mathrm{True})^{|V|}\). Observe that the tuple element-wise comparison at the right hand side is practically element-wise comparison of the columns of the distance matrix. This mean, we can systematically arrange all possible element-wise column comparison of the distance matrix into a new structure, called the distance similarity matrix,
	\[
		B^{(V)} = \left[\enspace D^{(V)}(\cdot, v_i)\ \ddot{=}\ D^{(V)}(\cdot, v_j) \enspace \cdots\enspace \right],\quad 1 \leq i < j \leq |V|
	\]
	which has a dimension of \(|V| \times \binom{|V|}{2}\). As a result, by construction, we are able to transform the resolving set criteria into a case of showing that no columns in this matrix contains all \(\mathrm{True}\) values, i.e. \(\forall C \overset{\mathrm{col}}{\in} B^{(V)}, C \neq (\mathrm{True})^{|V|}\). In addition, the distance similarity matrix preserves the row correspondence that \(V\) has with the original distance matrix, allowing for \(\forall u, v \in V, u \neq v, \exists C \overset{\mathrm{col}}{\in} B^{(V)}, C(W) = (r(u|V)\ \ddot{=}\ r(v|V))(W)\) to apply for any \(W \subseteq V\).

	...\textit{blah blah blah}...

	\section{...\textit{blah blah blah}...}

	...\textit{blah blah blah}...

	\section{Experiments and Results}

	...\textit{blah blah blah}...

	\section{Conclusions and Future Works}

	...\textit{blah blah blah}...
\end{document}
